`make.events`

### 2023-11-21 inst/gitbook/01_getting_started.Rmd
added functions quick overviews

Note however, that the printing information is now updated to state that you've added a `"modifiers"` object (even though it's the same as the default).

For more interesting simulations however, you can provide modifiers that actually modify the birth-death process.
We can create one for example that makes species go extinct if their ancestor has a negative trait value.
For that we need to create a `"modifiers"` object that modifies the `speciation` process with a specific condition and a specific modification when that condition is met.
For a speciation to occur (and a species to not go extinct), the algorithm draws a random value _x_ between 0 and 1 and if the value is smaller than the speciation parameter divided by the speciation + extinction parameter, the lineage speciates ($x < (\lambda \ (\lambda + \mu))$), else it goes extinct.

Here we can create a modifier forcing the lineage to go extinct every time a specific condition is met, e.g. if the lineage trait value is negative.
In other words, if the lineage's trait value is negative, make sure that _x_ is equal to 1, so that the speciation condition $x < (\lambda \ (\lambda + \mu))$ is never met.

@@

We can then provide these two functions (the condition `negative.ancestor` and how to modify the speciation event when this condition is met `going.extinct`).
If you are an advances `treats` user, you can design your own `speciation` function but if you just want to use a normal `speciation` function, you can use the default one from `treats` called `speciation`.

>>>>

Note however, that the printing information is now updated to state that you've added a `"modifiers"` object (even though it's the same as the default).
In this specific case the `modifiers` object is not necessary as it is doing nothing different than when using the default one (i.e. when no `modifiers` is provided).

For more interesting simulations however, you can provide modifiers that actually modify the birth-death process.
We can create one for example that makes species go extinct if their ancestor has a negative trait value.
For that we need to create a `"modifiers"` object that modifies the `speciation` process with a specific condition and a specific modification when that condition is met.
For a speciation to occur (and a species to not go extinct), the algorithm draws a random value _x_ between 0 and 1 and if the value is smaller than the speciation parameter divided by the speciation + extinction parameter, the lineage speciates ($x < (\lambda \ (\lambda + \mu))$), else it goes extinct.
_x_ is a random value that simulates some stochasticity in the evolutionary simulation.
At every step of the simulation, a different random value _x_ is drawn making sure that the resulting speciation or extinction is not predictable.

However, we can create a modifier that introduces less randomness (or a different level of randomness) by forcing the lineage to go extinct every time a specific condition is met, e.g. if the lineage trait value is negative.
This can be done for example by not randomly generating the value _x_ but giving it a fixed value of say 1 so that the estimation of the speciation function $x < (\lambda \ (\lambda + \mu))$ is always met.
In other words, if the lineage's trait value is negative, make sure that _x_ is equal to 1, so that the speciation condition $x < (\lambda \ (\lambda + \mu))$ is never met.

@@

This `going.extinct` function is to be contrasted with the normal modifier for selecting the value `x` which is `runif(1)` (drawing a random value between 0 and 1).
We can then provide these two functions (the condition `negative.ancestor` and how to modify the speciation event when this condition is met `going.extinct`).
If you are an advances `treats` user, you can design your own `speciation` function but if you just want to use a normal `speciation` function, you can use the default one from `treats` called `speciation`.

### 2023-11-21 inst/gitbook/02_make.traits.Rmd
Updated gitbook for v0.2
added functions quick overviews

@@

### Multidimensional traits and correlations

>>>>

### Quick overview

function | arguments | input | what does it do
---------|-----------|-------|----------------
`make.traits` | process* | a function | simulates the trait (see below)
 | n | a numeric value | the number of dimensions
 | start | one or more numeric values | the starting values for each dimensions
 | process.args | ... | anything to be passed to `process` (see below)
 | trait.names | a character string | the name(s) of the trait(s)
 | add | a `treats` object | another trait generated by `make.traits`
 | update | a `treats` object | another trait generated by `make.traits`
 | test | logical | whether to test the validity of the trait
 | background | a `treats` object | another trait generated by `make.traits` (to run in the background)
---------|-----------|-------|----------------
`process` | x0 | a numeric value | the value of the previous trait
 | edge.length | a numeric value | the branch length value
 | ... | ... | any other argument to be handled by the process

\* non-optional arguments

@@

## Multidimensional traits and correlations

### 2023-11-21 inst/gitbook/03_make.modifiers.Rmd
added functions quick overviews

>>>>

### Quick overview

function | arguments | input | what does it do
---------|-----------|-------|----------------
`make.modifiers` | branch.length | a function | generating branch length (see below)
 | selection | a function | selecting a lineage (see below)
 | speciation | a function | choosing between speciation and extinction (see below)
 | condition | a function | a function that affects the condition of when to trigger the functions above (see below)
 | modify | a function | a function that modifies the value of the trigger value generated when the `condition` is met (see below)
 | add | a `treats` object | another modifier generated by `make.modifiers`
 | update | a `treats` object | another modifier generated by `make.modifiers`
 | test | logical | whether to test the validity of the modifier
---------|-----------|-------|----------------
`branch.length`, `selection`, `speciation`, `condition` | bd.params | a `treats` object generated by `make.bd.params` | usually handled internally (contains the either the birth-death values or their distribution)
 | lineage | a named "lineage" list | usually handled internally (contains the current state of the simulations)
 | trait.values | a matrix | usually handled internally (contains the current generated traits of the simulations)

### 2023-11-21 inst/gitbook/04_make.events.Rmd
added functions quick overviews

>>>>

### Quick overview

  condition,
  modification,
  add,
  test = TRUE,
  event.name,
  replications = 0,
  additional.args
)

function | arguments | input | what does it do
---------|-----------|-------|----------------
`make.events` | target* | any of the following: "taxa", "bd.params", "traits", "modifiers" or "founding" | what aspect of the simulation should the event target
 | condition* | a function | a function that affects the condition of when to trigger the event (see below)
 | modification* | a function | a function that applies the modification to the target if the condition is triggered (see below)
 | add | a `treats` object | another event generated by `make.events`
 | test | logical | whether to test the validity of the event
 | event.name | a character string | the name of the event
 | replications | a number | if possible how many times is the event allowing to be repeated
 | additional.args | ... | any additional arguments to be passed to the event

Here is a list of implemented `conditions` (see `?events.conditions`):

function | arguments | input | what does it do
---------|-----------|-------|----------------
`time.condition`, `taxa.condition`, `trait.condition` | x | the value to check the target against | e.g. time = x, taxa = x or trait = x (see below)
 | condition | a relational operator | e.g. `==`, `>`, `!=`, etc. for asking for example `time == x` 
 | ... | any additional arguments to be passed to the event | |

Here is a list of implemented `modifications` (see `?events.modifications`):

 function | arguments | input | what does it do
---------|-----------|-------|----------------
random.extinction | x | a numeric value | the proportion of species to make extinct
trait.extinction | x | a numeric value | the trait value when to trigger the extinction
 | condition | a relational operator | the relation to the trait value (e.g. `x = 1` and `condition =  >` makes species that have a trait value > 1 go extinct)
bd.params.update | ... | any argument to be passed to `make.bd.params` | |
traits.update | ... | any argument to be passed to `make.traits` | |
modifiers.update | ... | any argument to be passed to `make.modifiers` | |

\* non-optional arguments
